Probing heavy ion collisions using quark and gluon jet substructure

### 2018-02-27 qg_ML.tex
YT updated the draft.

Jet substructure measurements provide concrete and consistent physics information of the modification of jets in a heavy ion environment. From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, the community has recently established that both transverse and longitudinal momentum distributions inside jets are significantly modified. Since jet substructures %and jet fragmentation 
depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may significantly contribute to substructure modifications observed in experiments. It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}. With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the changes in quark and gluon jet fractions caused by jet-medium interactions. This then allows one to focus on studying how quark and gluon jets are differently modified and use them as independent jet quenching probes. Note that the above jet modification picture implies that, with the high-purity quark jets in photon-tagged jet samples, the narrowing effect of jet cores should decrease. On the other hand, certain jet quenching models predict a universal narrowing of jet cores \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm,Casalderrey-Solana:2016jvj,Brewer:2017fqy} and emphasize contributions from medium responses \cite{Tachibana:2017syd}. Selecting purer samples of quark and gluon jets will provide the opportunity for directly studying how the medium responds to different hard probes.

@@

Due of the huge underlying event background in heavy ion collisions, jets with smaller radii are more reliably reconstructed and theoretically well understood \cite{Dasgupta:2014yra,Chien:2015cka,Becher:2015hka,Kang:2016mcy}. In order to minimize the background contamination within the experimentally allowed angular resolution, we study jets reconstructed using the anti-$k_T$ algorithm \cite{Cacciari:2008gp} with $R=0.4$. Within the catchment area of such jets, the significant background contamination needs to be subtracted before meaningful jet observable measurements can be made. In order to study multiple jet substructure observables and their correlations simultaneously, it necessarily requires a constituent-level, observable-independent background subtraction method \cite{Soyez:2012hv,Cacciari:2014gra,Berta:2014eza,Bertolini:2014bba,Komiske:2017ubm}. Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlights contributions from hard, collinear particles to jet observables. We will draw connections from telescoping deconstruction to the soft-drop groomed momentum sharing $z_g$ and groomed jet radius $r_g$ variables \cite{Larkoski:2014wba} and discuss qualitatively new jet substructure observables within the context of the telescoping deconstruction framework. We will also use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent and showcase significant regions of phase space in the context of the soft drop grooming procedure.

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
	   \caption{Distributions of invariant mass difference between soft-drop groomed mass and ungroomed mass for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
\label{fig:comp_delta_m2}
\end{figure}

Using the general idea of telescoping jets \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop jets: $\delta m^2 \equiv m^2_{\rm jet}-m^2_{\rm g-jet}$. Since soft-drop removes soft, wide angle radiation from jets, $\delta m^2$ gives another way of disentangling hard, collinear radiation and probe the soft contribution to jet mass. Figure~\ref{fig:comp_delta_m2} shows the $\delta m^2$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}. We see that $\delta m^2$ in pp collisions exhibits differences between quark jets and gluon jets. On the other hand, the $\delta m^2$ distributions for quark and gluon jets become very similar in PbPb collisions. This confirms again the previous observation that the soft event activities in \textsc{Jewel} washes out differences between quark and gluon jets so that the classification performance drops.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
	\caption{Similar to Fig:~\ref{fig:ROC_pp} ROC curves of quark v.s. gluon efficiency for particle jets in central PbPb (right panel) collisions in \textsc{Jewel}.}
\label{fig:ROC_pbpb}
\end{figure}

Figures~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb} show ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp and PbPb collisions respectively for particle jets, i.e. jets reconstructed with particle momenta without $\eta-\phi$ discretization and background subtraction. We see a remarkable overall compatibility between the methods with similar classification performances, implying that the three jet representations capture comparable jet information. However, there are noticeable performance differences especially in their ordering. Note the monotonic increase of the telescoping deconstruction performances from $\rm T1$ to $\rm T1+T2$, to $\rm T1+T2+T3$ with increasing amount of input information of higher-order subleading subjets. It is important to note that the TD classification performance saturates quickly at the T2 order, with additional higher-order features offering little improvement in the quark efficiency given a gluon efficiency. In pp collisions (Fig:~\ref{fig:ROC_pp}) DCNN outperforms the multivariate analysis using physics-motivated variables and has a similar performance as $\rm T1+T2+T3$. This implies that in the pp environment the $\eta-\phi$ discretization doesnt smear the relevant jet information, and the TD performance quickly saturates to the DCNN performance. In PbPb collisions (Fig:~\ref{fig:ROC_pbpb}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jwpy-simulated PbPb events. Measuring the particle multiplicity in heavy ion collisions is highly nontrivial due to the huge underlying event contribution and its correlation to jets. To take into account the difficulty of such an observable, we also show the ROC curve (the dotted green line in Fig:~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb}) for the MLP with physics-motivated variables exclusing the pixel multiplicity, which reduces the auc by a few percent $0.01 - 0.02$. Even though the TD with radius scan in steps of 0.08 is not able to accurately capture the full multiplicity information at the T3 order, it is still within a few percent difference amongst other methods in the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation. %Comparing Fig:~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb}, we see a stark degradation in the classification performance in PbPb collisions which we shall proceed to study in detail.

\begin{figure}[h]
	\centering
        \includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
	\caption{ROC curves of quark v.s. gluon efficiency using TD for grid jets in pp (dotted lines) and central PbPb (solid lines) collisions with GridSub subtraction. The different colors represent different TD orders (T1: blue, T2: red, T3: green). The corresponding area under the curves are shown in the brackets.}
	\label{fig:ROC_TD_grid}
\end{figure}

@@

In order to have a fair comparison between jets in pp and PbPb collisions (and to later on directly compare with data), we apply the grid discretization for jets in both pp and PbPb collisions, the later with the specific GridSub background subtraction technique, and compare with various TD orders in Fig~\ref{fig:ROC_TD_grid}. We clearly see the performance drops in PbPb collisions, along with a significant reduction in the performance improvement from T1 to $\rm T1+T2$ in PbPb collisions. This suggests that the information carried in higher-order subleading subjets can be washed out in \jw heavy ion collisions due to the medium interactions. In Fig~\ref{fig:ROC_TD3_grid} we plot the ROC curves for TD at third order and compare jets with or without $\eta-\phi$ discretization and background subtraction. Since the radius scan in TD respects the $\eta-\phi$ discretization resolution, we don't see the change of the TD performance using particle jets or grid jets. On the other hand, the performance drops very slightly (change in auc of $0.01$) with the background subtraction performed. This suggests some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction.

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.59\textwidth]{plots/JEWELPYTHIA_2p76TeV_quark_ppvspbpb.pdf}
	   \includegraphics[width=0.59\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
	   \caption{ROC curves of pp-jet v.s. PbPb-jet efficiency for quark (top panel) and gluon (left panel) jets with grid discretization. The different colored curves correspond to various classification models studied in this paper.}
%As before, the MLP with physics-motivated variables with and without pixel multiplicity in the green solid and dotted lines, DCNN on jet images in solid blue, TD at T1, T1+T2 and third order are in solid maroon, orange and cyan lines.}
\label{fig:ROC_qq_gg}
\end{figure}

@@

From our comprehensive comparisons of quark vs gluon jets, as well as comparing vacuum vs quenched jets, we can qualitatively understand the effects of quenching as implemented in the \jw Monte Carlo. Quenched jets end up with significantly more jet constituents which are distributed away from the jet axis, resulting in the saturation of the TD classification performance at the T1 order. This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can interact with a larger number of scattering centers leading to broader and more quenched jets. In the next section we will employ the use of another jet representation, the Lund Diagram, to illuminate the differences between quark and gluon jets, as well as jets in pp and PbPb collisions

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \jw. See text for the definition of the axes.}
\label{fig:Lund_full}
\end{figure}

Each jet observable is defined as a function of the momenta and angles between jet axis/constituents so as to probe the entire jet formation/structure with a single number. One can choose a set of jet observables to be a representation of jets, and we discussed previously, several approaches all faithfully encode the jet information. In this section we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation as it was recently shown to be efficient in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}. As we will see, the Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches. The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction. The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling limit for jets in pp collisions due to the infrared structure of QCD.

Given a jet reconstructed using the anti-$k_{T}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
\begin{itemize}
	\item Recluster the jet using the C/A algorithm
	\item Start from the root of the tree and move along the hard branches. At each branching, construct the momentum fraction of the soft branch $z=\min(p_{T_1},p_{T_2})/(p_{T_1}+p_{T_2})$ and the angular distance between the two branches $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\phi^2_{1, 2}}$
	\item Fill in the Lund diagram for each branching
	\item Continue along the subsequent hard branches until reaching only one particle
\end{itemize}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \jw. See text for the definition of the axes.}
\label{fig:Lund_bkg}
\end{figure}

In the Lund diagram, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope). Figure:~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions. We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets. Comparing quark to gluon jets, we see that gluon jets (right panel) end up with a larger fraction of the soft radiation. On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets. As we discussed in Fig:~\ref{fig:comp_z}, \ref{fig:comp_rg} and Fig:~\ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta. The main advantage of the Lund diagram is that one can directly see the emergence of such qualitative features.

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \jw. See text for the definition of the axes.}
\label{fig:Lund_hard}
\end{figure}

We can also study the effect of soft-drop using the Lund diagrams. Figure:~\ref{fig:Lund_bkg} shows the Lund diagrams for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables. That is, Fig:~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure. Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams. We see that a large portion of soft, wide-angle radiation is isolated from the hard, collinear core. The bottom right panel of Fig:~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation. These soft jet constituents near the jet boundary are precisely the ones that contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}. Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Fig:~\ref{fig:Lund_hard} shows the corresponding Lund diagram. We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets. However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions as shown by the enhanced wide angle radiation, even though now the angular scale has significantly reduced. This can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.

@@

We presented a systematic study of modifications to quark and gluon jet structure in heavy ion collisions, taking the \jwpy Monte Carlo simulations as an example. Jet modifications are studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets in different collision systems. We compare the performances of a MLP with physics-motivated jet variables, a state-of-the-art DCNN trained on jet images, and a newly developed method of telescoping deconstruction. We find that the quark gluon classification performance worsens for quenched jets, with a significant fraction of quark jets now masquerading as gluon jets. All the methods studied perform consistently, suggesting that the TD can extract fundamental jet fragmentation patterns. Through multiple methods and observables, we observe consistently that the dominant feature in the \jw PbPb collisions is the increase of soft particle multiplicity due to medium recoils throughout the jet region. This results in the lost of information in subleading subjets which is a characteristic feature of the \jw simulation.

The future of heavy ion jet physics program necessarily moves toward precise, simultaneous understanding of multiple jet substructure observables and their correlations. Therefore our work serves as an example how multivariate techniques can help illuminate the underlying mechanism of jet modifications. The telescoping deconstruction method suggests a framework of jet substructure observables for jet physics studies, which we show to give qualitatively new insights order-by-order in its observable organization. In the future, comparisons between theoretical calculations, simulations and experimental measurements will become standard practice to identify robust features of jet-medium interactions. Our work suggests a systematically improvable framework which will serve this important purpose.

\section*{Acknowledgments}
The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributions to the Appendix. The authors also thank Fr\'ed\'eric Dreyer, Benjamin Nachman, Joern Putschke, Konrad Tywoniuk and Sevil Salur for helpful discussions. We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications. RKE also thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques. YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008. RKE acknowledges supported in part by the U.S. Department of Energy Office of Science, Office of Nuclear Physics under Award Number DE-FG02-92ER-40713.

>>>>

Jet substructure measurements provide concrete and consistent physics information of the modification of jets in a heavy ion environment. From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, the community has recently established that both transverse and longitudinal momentum distributions inside jets are significantly modified. Since jet substructures %and jet fragmentation 
depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2012ur,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may significantly contribute to substructure modifications observed in experiments. It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}. With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the changes in quark and gluon jet fractions caused by jet-medium interactions. This then allows one to focus on studying how quark and gluon jets are differently modified and use them as independent jet quenching probes. Note that the above jet modification picture implies that, with the high-purity quark jets in photon-tagged jet samples, the narrowing effect of jet cores should decrease. On the other hand, certain jet quenching models predict a universal narrowing of jet cores \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm,Casalderrey-Solana:2016jvj,Brewer:2017fqy} and emphasize contributions from medium responses \cite{Tachibana:2017syd}. Selecting purer samples of quark and gluon jets will provide the opportunity for directly studying how the medium responds to different hard probes.

@@

Due of the huge underlying event background in heavy ion collisions, jets with smaller radii are more reliably reconstructed and theoretically well understood \cite{Dasgupta:2014yra,Chien:2015cka,Becher:2015hka,Kang:2016mcy}. In order to minimize the background contamination within the experimentally allowed angular resolution, we study jets reconstructed using the anti-$k_t$ algorithm \cite{Cacciari:2008gp} with $R=0.4$. Within the catchment area of such jets, the significant background contamination needs to be subtracted before meaningful jet observable measurements can be made. In order to study multiple jet substructure observables and their correlations simultaneously, it necessarily requires a constituent-level, observable-independent background subtraction method \cite{Soyez:2012hv,Cacciari:2014gra,Berta:2014eza,Bertolini:2014bba,Komiske:2017ubm}. Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlights contributions from hard, collinear particles to jet observables. We will draw connections from telescoping deconstruction to the soft-drop groomed momentum sharing $z_g$ and groomed jet radius $r_g$ variables \cite{Larkoski:2014wba} and discuss qualitatively new jet substructure observables within the context of the telescoping deconstruction framework. We will also use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent and showcase significant regions of phase space in the context of the soft drop grooming procedure.

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
	   \caption{Distributions of $\sqrt{\delta m^2}=\sqrt{m^2-m^2_{\rm soft~drop}}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
\label{fig:comp_delta_m2}
\end{figure}

Using the general idea of telescoping jet substructure \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop jets: $\delta m^2 \equiv m^2-m^2_{\rm soft~drop}$. Since soft-drop removes soft, wide angle radiation from jets, $\delta m^2$ disentangles hard, collinear radiation and probes the soft radiation that is dropped. Figure~\ref{fig:comp_delta_m2} shows the $\sqrt{\delta m^2}$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions simulated in \textsc{Jewel}. We see that the observable $\sqrt{\delta m^2}$ in pp collisions exhibits differences between quark jets and gluon jets. On the other hand, the distributions for quark and gluon jets become very similar in PbPb collisions. This hints that soft event activities in \textsc{Jewel} can wash out differences between quark and gluon jets, which leads to the lower classification performance as we will show in the next section.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
	\caption{ROC curves of quark v.s. gluon efficiency for particle jets in central PbPb (right panel) collisions in \textsc{Jewel}, similar to Figure~\ref{fig:ROC_pp}.}
\label{fig:ROC_pbpb}
\end{figure}

Figures~\ref{fig:ROC_pp} and Figure~\ref{fig:ROC_pbpb} show ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp and PbPb collisions, respectively, for particle jets, i.e. jets reconstructed with particle momenta without $\eta-\phi$ discretization and background subtraction. We see a remarkable overall compatibility between the methods with similar classification performances, implying that the three jet representations capture comparable jet information. However, there are noticeable performance differences especially in their ordering. Note the monotonic increase of the telescoping deconstruction performances from $\rm T1$ to $\rm T1+T2$, to $\rm T1+T2+T3$ with increasing amount of input information of higher-order subleading subjets. It is important to note that the TD classification performance saturates quickly at the T2 order, with additional higher-order features offering little improvement in the quark efficiency given a gluon efficiency. In pp collisions (Figure~\ref{fig:ROC_pp}) DCNN outperforms the multivariate analysis using physics-motivated variables and has a similar performance as $\rm T1+T2+T3$. This implies that in the pp environment the $\eta-\phi$ discretization doesnt smear the relevant jet information, and the TD performance quickly saturates to the DCNN performance. In PbPb collisions (Figure~\ref{fig:ROC_pbpb}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jwpy-simulated PbPb events. Measuring the particle multiplicity in heavy ion collisions is highly nontrivial due to the huge underlying event contribution and its correlation to jets. To take into account the difficulty of such an observable, we also show the ROC curve (the dotted green line in Figure~\ref{fig:ROC_pp} and Figure~\ref{fig:ROC_pbpb}) for the MLP with physics-motivated variables exclusing the pixel multiplicity, which reduces the auc by a few percent $0.01 - 0.02$. Even though the TD with radius scan in steps of 0.08 is not able to accurately capture the full multiplicity information at the T3 order, it is still within a few percent difference among other methods in the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation. %Comparing Fig:~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb}, we see a stark degradation in the classification performance in PbPb collisions which we shall proceed to study in detail.

\begin{figure}[h]
	\centering
        \includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
	\caption{ROC curves of quark v.s. gluon efficiency using TD for grid jets in pp (dotted lines) and central PbPb (solid lines) collisions in \textsc{Jewel} with GridSub subtraction. The different colors represent different TD orders (T1: blue, T2: red, T3: green). The corresponding area under the curves are shown in the brackets.}
	\label{fig:ROC_TD_grid}
\end{figure}

@@

In order to have a fair comparison between jets in pp and PbPb collisions (and to later on directly compare with data), we apply the grid discretization for jets in both pp and PbPb collisions, the later with the specific GridSub background subtraction technique, and compare with various TD orders in Figure~\ref{fig:ROC_TD_grid}. We clearly see the performance drops in PbPb collisions, along with a significant reduction in the performance improvement from T1 to $\rm T1+T2$ in PbPb collisions. This suggests that the information carried in higher-order subleading subjets can be washed out in \jw heavy ion collisions due to the medium interactions. In Figure~\ref{fig:ROC_TD3_grid} we plot the ROC curves for TD at the T3 order and compare jets with or without $\eta-\phi$ discretization and background subtraction. Since the radius scan in TD respects the $\eta-\phi$ discretization resolution, we don't see the change of the TD performance using particle jets or grid jets. On the other hand, the performance drops very slightly (change in auc of $0.01$) with the background subtraction performed. This suggests that some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction.

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.59\textwidth]{plots/JEWELPYTHIA_2p76TeV_quark_ppvspbpb.pdf}
	   \includegraphics[width=0.59\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
	   \caption{ROC curves of pp-jet v.s. PbPb-jet efficiency for quark (top panel) and gluon (bottom panel) jets with grid discretization. The different colored curves correspond to various classification models studied in this paper.}
%As before, the MLP with physics-motivated variables with and without pixel multiplicity in the green solid and dotted lines, DCNN on jet images in solid blue, TD at T1, T1+T2 and third order are in solid maroon, orange and cyan lines.}
\label{fig:ROC_qq_gg}
\end{figure}

@@

From our comprehensive comparisons of quark v.s. gluon jets, as well as comparing vacuum v.s. quenched jets, we can qualitatively understand the effects of quenching as implemented in the \jw Monte Carlo. Quenched jets end up with significantly more jet constituents which are distributed away from the jet axis, resulting in the saturation of the TD classification performance at the T1 order. This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can interact with a larger number of scattering centers leading to broader and more quenched jets. In the next section we will employ the use of another jet representation, the Lund Diagram, to illuminate the differences between quark and gluon jets, as well as jets in pp and PbPb collisions

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
\label{fig:Lund_full}
\end{figure}

Each jet observable is defined as a function of the constituent momenta and angles from jet axes so as to probe the entire jet substructure with a single number. One can choose a set of jet observables to be a representation of jets, and we discussed previously several approaches all faithfully encode the jet information. In this section we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation. It was shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}. As we will see, the Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches. The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction. The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD.

Given a jet reconstructed using the anti-$k_{t}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
\begin{itemize}
	\item Recluster the jet using the C/A algorithm
	\item Start from the root of the tree and move along the hard branches. At each branching, construct the momentum fraction of the soft branch $z=\min(p_{T_1},p_{T_2})/(p_{T_1}+p_{T_2})$ and the angular distance between the two branches $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\phi^2_{1, 2}}$
	\item Fill in the Lund diagram for each branching
	\item Continue along the subsequent hard branches until reaching only one particle
\end{itemize}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
\label{fig:Lund_bkg}
\end{figure}

In the Lund diagram, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope). Figure~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions in \textsc{Jewel}. We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets. Comparing quark to gluon jets, we see that gluon jets (right panel) have a larger fraction of soft radiation. On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets. As we discussed in Figures~\ref{fig:comp_z}, \ref{fig:comp_rg} and \ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta. Using the Lund diagram one can directly see the emergence of such qualitative features.

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
\label{fig:Lund_hard}
\end{figure}

We can also study the effect of soft-drop using the Lund diagrams. Figure~\ref{fig:Lund_bkg} shows the Lund diagrams for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables. That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure. Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams. We see that a large portion of soft, wide-angle radiation is isolated from the hard, collinear core. The bottom right panel of Figure~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation. These soft jet constituents near the jet boundary are precisely the ones that contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}. Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Figure~\ref{fig:Lund_hard} shows the corresponding Lund diagram. We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets. However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions. A similar observation was made in \cite{Hoang:2017kmk} that hard branchings can ``protect" soft radiation from being dropped. This is shown by the enhanced, wide angle radiation, even though now its angular scale has significantly reduced. The radiation can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.

@@

We presented a systematic study of modifications to quark and gluon jet substructure in heavy ion collisions, taking the \jwpy Monte Carlo simulations as an example. Jet modifications are studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets in different collision systems. We compare the performances of a MLP with physics-motivated jet variables, a state-of-the-art DCNN trained on jet images, and a newly developed method of telescoping deconstruction. We find that the quark gluon classification performance worsens for quenched jets, with a significant fraction of quark jets now masquerading as gluon jets. All the methods studied perform consistently, suggesting that the TD can extract fundamental jet fragmentation patterns. Through multiple methods and observables, we observe consistently that the dominant feature in the \jw PbPb collisions is the increase of soft particle multiplicity due to medium recoils throughout the jet region. This is closely related to the lost of information in subleading subjets, which is a characteristic feature of the \jw simulation.

The future of heavy ion jet physics program necessarily moves toward precise, simultaneous understanding of multiple jet substructure observables and their correlations. Therefore our work serves as an example of how multivariate techniques can help illuminate the underlying mechanism of jet modifications. The telescoping deconstruction method suggests a framework of jet substructure observables for jet physics studies, which we show to give qualitatively new insights order-by-order in its observable organization. In the future, comparisons between theoretical calculations, simulations and experimental measurements will become standard practice to identify robust features of jet-medium interactions. Our work suggests a systematically improvable framework which will serve this important purpose.

\section*{Acknowledgments}
The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributions to the Appendix. The authors also thank Fr\'ed\'eric Dreyer, Benjamin Nachman, Joern Putschke, Jesse Thaler, Konrad Tywoniuk and Sevil Salur for helpful discussions. We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications. RKE also thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques. YTC also thank the organizers of the BOOST2015 conference where the idea of telescoping deconstruction was first presented. YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008. RKE acknowledges supports in part by the U.S. Department of Energy Office of Science, Office of Nuclear Physics under Award Number DE-FG02-92ER-40713.
